---
title: "705 Cox multi-sequence sieve analysis"
output:
  pdf_document: default
  html_notebook: default
---

# Introduction 

In this document, we implement a multi-sequence sieve analysis using the methods 
to analyze binary marks from Peng et al. (in progress). 

*Data processing notes*: 

* Of 2,188 participants, there are 119 with `hiv1event == 1`. 
* There are 23 participants with sequencing data for whom `hiv1event == 0`. 
* There are 3 participants for whom `hiv1event == 1` with no sequencing data. 
(These participants are removed for the time being.)
* With these 3 excluded, we have 2,185 participants in the dataset. 

*Covariates*: BMI, risk score, age, stratified by indicator of South Africa

*Pre-screening*: 

* For each binary mark, calculate the raw proportions of each individual's sequences
having mark value = 1. 
* Let $\tilde{J}_{i,q_0}$ denote the indicator that the $i$th individual's 
raw proportion exceeds the $q_0$ threshold. 
* Let $J_{i, mindist}$ denote the mindist sequence binary value of the $i$th 
individual. 
* If (1) $\sum_{i=1}^{n_{cases}} | J_{i, mindist} -\tilde{J}_{i,q_0}| \ge 12$ (over 10%
of threshold-classified marks are different than the mindist marks ), (2) 
$\sum_{i=1}^{n_{cases}} \tilde{J}_{i,q_0} \ge 4$, and (3) 
$\sum_{i=1}^{n_{cases}} |1 - \tilde{J}_{i,q_0}| \ge 4$, then we screen in the mark 
for the multi-sequence analysis. 

# Setup 

* Make sure that the multi-sequence sieve dataset, mindist sieve dataset, and 
subject-level sieve dataset are in the same folder as this code file, or change 
the directory names in the data section below. 
* Make sure all packages are installed. 
* For each binary mark type, this file will output CSV files with the pre-screening 
metrics and the multi-sequence sieve analysis results for screened in results. 

# Packages 

```{r, include = F}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = F)
```


```{r include = T, warning = F, message = F}
library(dplyr)
library(DT)
library(purrr)
library(ggplot2)
library(ggpubr)
library(survival)
library(deconvolveR)
library(future)
library(future.apply)
library(ggbeeswarm)

source("helper/multiseq_cox_code.R")
source("helper/deconv_fast.R")
source("helper/p.adj.perm2.R")

plan(multisession, workers = parallel::detectCores() - 1)

boots <- 300
```

# Read datasets 


```{r}
# multi-sequence sieve dataset 
df_sieve <- read.csv("HVTN_705_multiseq_sieve_data_v3.csv")

# mindist sieve dataset 
df_sieve_mindist <- read.csv("HVTN_705_sieve_data_final_v8.csv")

# subject-level sieve dataset 
df_sieve_subj <- read.csv("HVTN_705_multiseq_subject_data_v3.csv")

```

# Data processing 

```{r}
# merge in hiv1event and arm 
df_sieve <- df_sieve %>% 
  
  merge(df_sieve_subj %>% 
          select(subjid, hiv1event, armdesc))

# filter for first visit only 
df_sieve_visit_to_keep <- df_sieve %>% 
  arrange(subjid, visit) %>% 
  select(subjid, visit) %>% 
  distinct(subjid, .keep_all = T)

df_sieve <- df_sieve %>% merge(df_sieve_visit_to_keep)

```

# Residue matched to insert {.tabset .tabset-fade} 

"hxb2.pos.1mer.insert.match[.tier1]": The binary indicator variable for 
whether a residue at the given site matches the residue in the indicated 
insert.

* "pos" = the HXB2 coordinate of the position 
* "insert" = the insert that the representative sequence is compared with.

Our inserts are coded as:

* "mos1" = Ad26 Mos1 Env insert
* "mos2" = Ad26 Mos2S Env insert
* "c97za" = C97ZA Clade C insert
* "1428v1v2" = V1V2 positions motivated by the V1V2 IgG3 correlates result,
                   compared to the 1428 V1V2 antigen sequence.

The "tier1" tag is present only for those sites that are classified as Tier 1,
Type 2 marks.

Matches for "mos1", "mos2", and "c97za" were made for all positions in Env
which were not deemed unalignable.  This includes positions that did not
pass the minimum variability filter.

Matches for "1428v1v2" were made for positions 130, 155, 160, 161, 164, 165,
166, 170, 172, 173, 178, 179, 181, 186, and 200.
   
A value of "1" indicates a match with the vaccine, while a value of "0" 
indicates a mismatch.

```{r}
cols_to_summarize_match <- grep("^hxb2\\.[0-9]+\\.1mer\\..+$", colnames(df_sieve), 
                                value = TRUE)
```

## $q_0 = 0.01$  

### Variability screen 

Outputs datasets with the following variables: 

* prop_other: Proportion of observations where the raw mark value 
`Q = K / sequence_depth` is strictly between 0 and 1.
* cat0_sum: Number of observations classified as $J_{q0} = 0$ (raw mark below the 
threshold).
* cat1_sum: Number of observations classified as $J_{q0} = 1$ (raw mark at or 
above the threshold).
* mindist0_sum: Number of observations where the mindist-based mark equals 1.
* mindist1_sum: Number of observations where the mindist-based mark equals 0.
diff_from_mindist: Sum of absolute differences between the raw classification 
$J_q0$ and the transformed mindist mark `1 - mindist_mark`.
* pass_screen: Indicator of whether the dataset meets the variability criteria 

```{r}
df_prescreening_match_0.01 <- data.frame()

for (col in cols_to_summarize_match){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_variability_screen(df_mark,
                                     thresh = 0.01)

  res_temp$mark <- col

  df_prescreening_match_0.01 <- rbind(df_prescreening_match_0.01, res_temp)
}

df_prescreening_match_0.01 <- df_prescreening_match_0.01 %>%
  mutate(pass_screen = ifelse(diff_from_mindist >= 12 & cat0_sum >= 4 &
                                cat1_sum >= 4, TRUE, FALSE))

write.csv(df_prescreening_match_0.01,
          file = "prescreening_match_0.01_screen.csv",
          row.names = F)
```

### Sieve analysis 

Outputs datasets with the following variables: 

* ve0_estimate: VE estimate for $J_{q_0} = 0$ viruses.
* ve0_se: Bootstrap standard error for VE0.
* ve0_lo: Lower confidence limit for VE0.
* ve0_hi: Upper confidence limit for VE0.
* ve1_estimate: VE estimate for J_q0 = 1 viruses.
* ve1_se: Bootstrap standard error for VE1.
* ve1_lo: Lower confidence limit for VE1.
* ve1_hi: Upper confidence limit for VE1.
* p_ve0: P-value for testing VE0 = 0.
* p_ve1: P-value for testing VE1 = 0.
* p_val_no_ve: P-value for testing that VE0 = VE1 = 0.
* p_val_equal: P-value for testing that VE0 = VE1.

```{r}
temp_filtered_cols <- (df_prescreening_match_0.01 %>% filter(pass_screen == T))$mark
thresh <- 0.01 

df_res <- data.frame()
for (col in temp_filtered_cols){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_full_analysis(df_mark,
                                cl_probs_vars = c("Z"),
                                covariates = c("Z", "age", "bmi", "riskscoresl"),
                                event_var = "allstatus",
                                time_var = "observed_time",
                                stratification_var = "ind_sa",
                                boots = boots,
                                thresh = thresh,
                                par = rep(0.5, 8),
                                col_beta0 = 1,
                                col_beta1 = 5)

  res_temp$mark <- col
  res_temp$thresh <- thresh

  df_res <- df_res %>%
    rbind(res_temp)
}

write.csv(df_res,
          file = "res_match_0.01.csv",
          row.names = F)
```


## $q_0 = 0.99$  

### Variability screen 

Outputs datasets with the following variables: 

* prop_other: Proportion of observations where the raw mark value 
`Q = K / sequence_depth` is strictly between 0 and 1.
* cat0_sum: Number of observations classified as $J_{q0} = 0$ (raw mark below the 
threshold).
* cat1_sum: Number of observations classified as $J_{q0} = 1$ (raw mark at or 
above the threshold).
* mindist0_sum: Number of observations where the mindist-based mark equals 1.
* mindist1_sum: Number of observations where the mindist-based mark equals 0.
diff_from_mindist: Sum of absolute differences between the raw classification 
$J_q0$ and the transformed mindist mark `1 - mindist_mark`.
* pass_screen: Indicator of whether the dataset meets the variability criteria 

```{r}
df_prescreening_match_0.99 <- data.frame()

for (col in cols_to_summarize_match){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_variability_screen(df_mark,
                                     thresh = 0.99)

  res_temp$mark <- col

  df_prescreening_match_0.99 <- rbind(df_prescreening_match_0.99, res_temp)
}

df_prescreening_match_0.99 <- df_prescreening_match_0.99 %>%
  mutate(pass_screen = ifelse(diff_from_mindist >= 12 & cat0_sum >= 4 &
                                cat1_sum >= 4, TRUE, FALSE))

write.csv(df_prescreening_match_0.99,
          file = "prescreening_match_0.99_screen.csv",
          row.names = F)

```



### Sieve analysis 

Outputs datasets with the following variables: 

* ve0_estimate: VE estimate for $J_{q_0} = 0$ viruses.
* ve0_se: Bootstrap standard error for VE0.
* ve0_lo: Lower confidence limit for VE0.
* ve0_hi: Upper confidence limit for VE0.
* ve1_estimate: VE estimate for J_q0 = 1 viruses.
* ve1_se: Bootstrap standard error for VE1.
* ve1_lo: Lower confidence limit for VE1.
* ve1_hi: Upper confidence limit for VE1.
* p_ve0: P-value for testing VE0 = 0.
* p_ve1: P-value for testing VE1 = 0.
* p_val_no_ve: P-value for testing that VE0 = VE1 = 0.
* p_val_equal: P-value for testing that VE0 = VE1.

```{r}
temp_filtered_cols <- (df_prescreening_match_0.99 %>% filter(pass_screen == T))$mark

thresh <- 0.99

df_res <- data.frame()
for (col in temp_filtered_cols){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_full_analysis(df_mark,
                                cl_probs_vars = c("Z"),
                                covariates = c("Z", "age", "bmi", "riskscoresl"),
                                event_var = "allstatus",
                                time_var = "observed_time",
                                stratification_var = "ind_sa",
                                boots = boots,
                                thresh = thresh,
                                par = rep(0.5, 8),
                                col_beta0 = 1,
                                col_beta1 = 5)

  res_temp$mark <- col
  res_temp$thresh <- thresh

  df_res <- df_res %>%
    rbind(res_temp)
}

write.csv(df_res,
          file = "res_match_0.99.csv",
          row.names = F)

```


# Reference-independent amino acid scanning {.tabset .tabset-fade} 

hxb2.pos.is.aa: The binary indictor variable for the reference-independent
amino acid site scanning.

* "pos" = the HXB2 coordinate of the position. Note that the values for
              "pos" have leading zeros for positions that are less than 100.
* "aa" = the amino acid that exists at that position for at least one
             sequence. If there is a gap at that position, the value of "aa"
             will be "gap".

A value of "1" indicates a match with the amino acid, while a value of "0"
indicates a mismatch.

```{r}
cols_to_summarize_aa <- grep("^hxb2\\..*\\.is\\..*$", colnames(df_sieve),
                             value = TRUE)
```

## $q_0 = 0.01$  

### Variability screen 

Outputs datasets with the following variables: 

* prop_other: Proportion of observations where the raw mark value 
`Q = K / sequence_depth` is strictly between 0 and 1.
* cat0_sum: Number of observations classified as $J_{q0} = 0$ (raw mark below the 
threshold).
* cat1_sum: Number of observations classified as $J_{q0} = 1$ (raw mark at or 
above the threshold).
* mindist0_sum: Number of observations where the mindist-based mark equals 1.
* mindist1_sum: Number of observations where the mindist-based mark equals 0.
diff_from_mindist: Sum of absolute differences between the raw classification 
$J_q0$ and the transformed mindist mark `1 - mindist_mark`.
* pass_screen: Indicator of whether the dataset meets the variability criteria 


```{r}
df_prescreening_aa_0.01 <- data.frame()

for (col in cols_to_summarize_aa){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_variability_screen(df_mark,
                                      thresh = 0.01)

  res_temp$mark <- col

  df_prescreening_aa_0.01 <- rbind(df_prescreening_aa_0.01, res_temp)
}

df_prescreening_aa_0.01 <- df_prescreening_aa_0.01 %>%
  mutate(pass_screen = ifelse(diff_from_mindist >= 12 & cat0_sum >= 4 &
                                cat1_sum >= 4, TRUE, FALSE))

write.csv(df_prescreening_aa_0.01,
          file = "prescreening_aa_0.01_screen.csv",
          row.names = F)
```


### Sieve analysis 

Outputs datasets with the following variables: 

* ve0_estimate: VE estimate for $J_{q_0} = 0$ viruses.
* ve0_se: Bootstrap standard error for VE0.
* ve0_lo: Lower confidence limit for VE0.
* ve0_hi: Upper confidence limit for VE0.
* ve1_estimate: VE estimate for J_q0 = 1 viruses.
* ve1_se: Bootstrap standard error for VE1.
* ve1_lo: Lower confidence limit for VE1.
* ve1_hi: Upper confidence limit for VE1.
* p_ve0: P-value for testing VE0 = 0.
* p_ve1: P-value for testing VE1 = 0.
* p_val_no_ve: P-value for testing that VE0 = VE1 = 0.
* p_val_equal: P-value for testing that VE0 = VE1.

```{r}
temp_filtered_cols <- (df_prescreening_aa_0.01 %>% filter(pass_screen == T))$mark

thresh <- 0.01
df_res <- data.frame()
for (col in temp_filtered_cols){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_full_analysis(df_mark,
                                cl_probs_vars = c("Z"),
                                covariates = c("Z", "age", "bmi", "riskscoresl"),
                                event_var = "allstatus",
                                time_var = "observed_time",
                                stratification_var = "ind_sa",
                                boots = boots,
                                thresh = thresh,
                                par = rep(0.5, 8),
                                col_beta0 = 1,
                                col_beta1 = 5)

  res_temp$mark <- col
  res_temp$thresh <- thresh

  df_res <- df_res %>%
    rbind(res_temp)

}

write.csv(df_res,
          file = "res_aa_0.01.csv",
          row.names = F)
```


## $q_0 = 0.99$  

### Variability screen 

Outputs datasets with the following variables: 

* prop_other: Proportion of observations where the raw mark value 
`Q = K / sequence_depth` is strictly between 0 and 1.
* cat0_sum: Number of observations classified as $J_{q0} = 0$ (raw mark below the 
threshold).
* cat1_sum: Number of observations classified as $J_{q0} = 1$ (raw mark at or 
above the threshold).
* mindist0_sum: Number of observations where the mindist-based mark equals 1.
* mindist1_sum: Number of observations where the mindist-based mark equals 0.
diff_from_mindist: Sum of absolute differences between the raw classification 
$J_q0$ and the transformed mindist mark `1 - mindist_mark`.
* pass_screen: Indicator of whether the dataset meets the variability criteria 

```{r}
df_prescreening_aa_0.99 <- data.frame()

for (col in cols_to_summarize_aa){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_variability_screen(df_mark,
                                     thresh = 0.99)

  res_temp$mark <- col

  df_prescreening_aa_0.99 <- rbind(df_prescreening_aa_0.99, res_temp)
}

df_prescreening_aa_0.99 <- df_prescreening_aa_0.99 %>%
  mutate(pass_screen = ifelse(diff_from_mindist >= 12 & cat0_sum >= 4 &
                                cat1_sum >= 4, TRUE, FALSE))

write.csv(df_prescreening_aa_0.99,
          file = "prescreening_aa_0.99_screen.csv",
          row.names = F)

```


### Sieve analysis 

Outputs datasets with the following variables: 

* ve0_estimate: VE estimate for $J_{q_0} = 0$ viruses.
* ve0_se: Bootstrap standard error for VE0.
* ve0_lo: Lower confidence limit for VE0.
* ve0_hi: Upper confidence limit for VE0.
* ve1_estimate: VE estimate for J_q0 = 1 viruses.
* ve1_se: Bootstrap standard error for VE1.
* ve1_lo: Lower confidence limit for VE1.
* ve1_hi: Upper confidence limit for VE1.
* p_ve0: P-value for testing VE0 = 0.
* p_ve1: P-value for testing VE1 = 0.
* p_val_no_ve: P-value for testing that VE0 = VE1 = 0.
* p_val_equal: P-value for testing that VE0 = VE1.

```{r}
temp_filtered_cols <- (df_prescreening_aa_0.99 %>% filter(pass_screen == T))$mark

thresh <- 0.99 
df_res <- data.frame()
for (col in temp_filtered_cols){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)

  res_temp <- run_full_analysis(df_mark,
                                cl_probs_vars = c("Z"),
                                covariates = c("Z", "age", "bmi", "riskscoresl"),
                                event_var = "allstatus",
                                time_var = "observed_time",
                                stratification_var = "ind_sa",
                                boots = boots,
                                thresh = thresh,
                                par = rep(0.5, 8),
                                col_beta0 = 1,
                                col_beta1 = 5)

  res_temp$mark <- col
  res_temp$thresh <- thresh

  df_res <- df_res %>%
    rbind(res_temp)
}

write.csv(df_res,
          file = "res_aa_0.99.csv",
          row.names = F)
```


# Presence of sequon {.tabset .tabset-fade} 

"hxb2.pos.is.sequon.tier1": Binary variables indicating the presence ("1")
or absence ("0") of a sequon (i.e., the canonical N-linked glycosylation 
motif:  [N][!P][S|T]) starting at site "pos".  
   
These marks were included for a total of 10 sites:  those associated with 
glycan holes in the C97ZA insert (sites 234 and 392), and those associated
with V2 bnAb contact glycans (sites 130, 156, 160, 234, 332, and 392).  Two 
of these sites are redundant (234 and 392), so this amounts to eight unique
sites total.

Unalignable sites were excluded, but all of these eight sites were alignable,
so they were all included.
   
All of these sites are Tier 1, Type 2 marks, so they all have the "tier1" 
tag.

For a sensitivity analysis, two new sites were included in this data 
category:  sites 362 and 363.  These address the sieve effect at site 364,
which appears to potentially disrupt sequons at these positions.

```{r}
cols_to_summarize_sequon <- grep(".is.sequon.tier1", colnames(df_sieve), 
                                 value = TRUE)
```

## $q_0 = 0.01$  

### Variability screen 

Outputs datasets with the following variables: 

* prop_other: Proportion of observations where the raw mark value 
`Q = K / sequence_depth` is strictly between 0 and 1.
* cat0_sum: Number of observations classified as $J_{q0} = 0$ (raw mark below the 
threshold).
* cat1_sum: Number of observations classified as $J_{q0} = 1$ (raw mark at or 
above the threshold).
* mindist0_sum: Number of observations where the mindist-based mark equals 1.
* mindist1_sum: Number of observations where the mindist-based mark equals 0.
diff_from_mindist: Sum of absolute differences between the raw classification 
$J_q0$ and the transformed mindist mark `1 - mindist_mark`.
* pass_screen: Indicator of whether the dataset meets the variability criteria 

```{r}
df_prescreening_sequon_0.01 <- data.frame()

for (col in cols_to_summarize_sequon){
  df_mark <- create_mark_dataset(df_sieve,
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)
  
  res_temp <- run_variability_screen(df_mark, 
                                     thresh = 0.01)
  
  res_temp$mark <- col
  
  df_prescreening_sequon_0.01 <- rbind(df_prescreening_sequon_0.01, res_temp)
}

df_prescreening_sequon_0.01 <- df_prescreening_sequon_0.01 %>%
  mutate(pass_screen = ifelse(diff_from_mindist >= 12 & cat0_sum >= 4 & 
                                cat1_sum >= 4, TRUE, FALSE))

write.csv(df_prescreening_sequon_0.01,
          file = "prescreening_sequon_0.01_screen.csv",
          row.names = F)

```


### Sieve analysis 

No screened in features. 

## $q_0 = 0.99$  

### Variability screen 

Outputs datasets with the following variables: 

* prop_other: Proportion of observations where the raw mark value 
`Q = K / sequence_depth` is strictly between 0 and 1.
* cat0_sum: Number of observations classified as $J_{q0} = 0$ (raw mark below the 
threshold).
* cat1_sum: Number of observations classified as $J_{q0} = 1$ (raw mark at or 
above the threshold).
* mindist0_sum: Number of observations where the mindist-based mark equals 1.
* mindist1_sum: Number of observations where the mindist-based mark equals 0.
diff_from_mindist: Sum of absolute differences between the raw classification 
$J_q0$ and the transformed mindist mark `1 - mindist_mark`.
* pass_screen: Indicator of whether the dataset meets the variability criteria 

```{r}
df_prescreening_sequon_0.99 <- data.frame()

for (col in cols_to_summarize_sequon){
  df_mark <- create_mark_dataset(df_sieve, 
                                 df_sieve_subj,
                                 df_sieve_mindist,
                                 mark = col)
  
  res_temp <- run_variability_screen(df_mark, 
                                     thresh = 0.99)
  
  res_temp$mark <- col
  
  df_prescreening_sequon_0.99 <- rbind(df_prescreening_sequon_0.99, res_temp)
}

df_prescreening_sequon_0.99 <- df_prescreening_sequon_0.99 %>%
  mutate(pass_screen = ifelse(diff_from_mindist >= 12 & cat0_sum >= 4 & 
                                cat1_sum >= 4, TRUE, FALSE))

write.csv(df_prescreening_sequon_0.99,
          file = "prescreening_sequon_0.99_screen.csv",
          row.names = F)

```

### Sieve analysis 

No screened in features. 

# Permutation test 

Code for running the permutation test. Note that this takes a very long time
to run! 

```{r}
# obtained screened-in marks 
df_analysis_params <- 
  (df_prescreening_match_0.01 %>% 
     filter(pass_screen == T) %>% 
     mutate(thresh = 0.01)) %>% 
  
  rbind(
    df_prescreening_match_0.99 %>% 
     filter(pass_screen == T) %>% 
     mutate(thresh = 0.99) 
  ) %>% 

  rbind(
    df_prescreening_aa_0.01 %>% 
     filter(pass_screen == T) %>% 
     mutate(thresh = 0.01)
  ) %>% 
  
  rbind(
    df_prescreening_aa_0.99 %>% 
     filter(pass_screen == T) %>% 
     mutate(thresh = 0.99)
  ) 

# obtain unadjusted p-values 
df_res1 <- read.csv("res_match_0.01.csv") %>% 
  select(mark, thresh, p_val_equal)

df_res2 <- read.csv("res_match_0.99.csv") %>% 
  select(mark, thresh, p_val_equal)

df_res3 <- read.csv("res_aa_0.01.csv")  %>% 
  select(mark, thresh, p_val_equal)

df_res4 <- read.csv("res_aa_0.99.csv")  %>% 
  select(mark, thresh, p_val_equal)

df_res_all <- df_res1 %>% 
  rbind(df_res2) %>% 
  rbind(df_res3) %>% 
  rbind(df_res4)


# select screened in marks in sieve dataset
df_sieve_select <- df_sieve %>% 
  select(c(subjid, df_analysis_params$mark))


# perform permutation test
nPerm <- 300

# permutated p-values 
pvals <- future_lapply(1:(nPerm + 1), function(i) {
  print(i)
  df_mark_dummy <- create_mark_dataset(df_sieve_select, 
                                       df_sieve_subj,
                                       df_sieve_mindist_select,
                                       mark = df_analysis_params[1, "mark"]) 
  
  idx <- sample.int(sum(df_mark_dummy$allstatus == 1))  
  
  res_vec <- numeric(nrow(df_analysis_params))

  # run sieve analysis for all setting specified in df_analysis_params 
  for (i in 1:nrow(df_analysis_params)) {
    temp_params <- df_analysis_params[i, ]
    
    df_mark <- create_mark_dataset(df_sieve_select, 
                                   df_sieve_subj,
                                   df_sieve_mindist_select,
                                   mark = temp_params$mark)
      
    df_mark$sequence_depth[df_mark$allstatus == 1] <- 
      df_mark$sequence_depth[df_mark$allstatus == 1][idx] 
    
    df_mark$K[df_mark$allstatus == 1] <- 
      df_mark$K[df_mark$allstatus == 1][idx]  
      
    res_temp <- run_full_analysis(df_mark,
                                  cl_probs_vars = c("Z"),
                                  covariates    = c("Z", "age", "bmi", "riskscoresl"),
                                  event_var     = "allstatus",
                                  time_var      = "observed_time",
                                  stratification_var = "ind_sa",
                                  boots         = 300,
                                  thresh        = temp_params$thresh,
                                  par           = rep(0.5, 8),
                                  col_beta0     = 1,
                                  col_beta1     = 5)
    res_vec[i] <- as.numeric(res_temp$p_val_equal)
  }
  
  res_vec
}, future.seed = TRUE)

pvals_mat <- do.call(rbind, pvals)

# run permutation test
ans <- p.adj.perm2(df_res_all$p_val_equal, permutation_mat)
write.csv(ans,
          file = "permutation_p_values.csv",
          row.names = F)
```


# Specific analyses - hxb2.832.is.L

## Figure 

The code below produces the plot of the raw proportions of hxb2.832.is.L. 

```{r}
df_mark <- create_mark_dataset(df_sieve,
                               df_sieve_subj,
                               df_sieve_mindist,
                               mark = "hxb2.832.is.L")

means <- df_mark %>% 
  mutate(raw_Q = (sequence_depth - K) / sequence_depth) %>% 
  filter(!is.na(raw_Q)) %>% 
  group_by(Z) %>% 
  summarise(mean_raw_Q = mean(raw_Q), .groups = "drop")  %>% 
  mutate( Z = ifelse(Z == 0, "Placebo", "Vaccine"))


p <- df_mark %>%
  mutate(
    raw_Q = (sequence_depth - K) / sequence_depth,
    Z = ifelse(Z == 0, "Placebo", "Vaccine"),
    y_order = rank(sequence_depth, ties.method = "first") 
  ) %>%
  filter(!is.na(raw_Q)) %>%
  ggplot(aes(x = raw_Q, y = y_order, color = Z, size = sqrt(sequence_depth))) +
  geom_quasirandom(
    width = 0,              # no horizontal spread
    groupOnX = FALSE,       # ensures jitter applied across all x values
    alpha = 0.8
  ) +
  geom_vline(xintercept = 0.99, linetype = "dashed", color = "red", size = 1.5) +
  geom_vline(data = means, aes(xintercept = mean_raw_Q, color = Z), size = 3) +
  facet_grid(rows = vars(Z), scales = "free_y") +
  scale_color_manual(values = c("Placebo" = "darkblue", "Vaccine" = "orange")) +
  xlab("Within-Host L832 Prevalence") +
  ylab(NULL) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(size = 30),
    axis.title = element_text(size = 30),
    strip.text = element_text(size = 30)
  )

ggsave(
  filename = "L832_prevalence_plot_gaps_included.pdf",
  plot = p,
  width = 16,       # inches
  height = 8,       # adjust as desired
  dpi = 600,        # publication-quality
  units = "in"
)
```


## Sensitivity analysis 

### Removing gaps 

The code below runs the analysis excluding gaps. 

```{r}
df_mark <- create_mark_dataset(df_sieve %>% 
                                 filter(hxb2.832.is.gap != 1 ),
                               df_sieve_subj,
                               df_sieve_mindist,
                               mark = "hxb2.832.is.L")

res_temp <- run_full_analysis(df_mark,
                              cl_probs_vars = c("Z"),
                              covariates = c("Z", "age", "bmi", "riskscoresl"),
                              event_var = "allstatus",
                              time_var = "observed_time",
                              stratification_var = "ind_sa",
                              boots = 300,
                              thresh = 0.01,
                              par = rep(0.5, 8),
                              col_beta0 = 1,
                              col_beta1 = 5)

res_temp
```

### Figure without gaps 

The code below produces the plot of the raw proportions of hxb2.832.is.L 
excluding gaps. 

```{r}
df_mark <- create_mark_dataset(df_sieve %>% 
                                 filter(hxb2.832.is.gap != 1),
                               df_sieve_subj,
                               df_sieve_mindist,
                               mark = "hxb2.832.is.L")

means <- df_mark %>% 
  mutate(raw_Q = (sequence_depth - K) / sequence_depth) %>% 
  filter(!is.na(raw_Q)) %>% 
  group_by(Z) %>% 
  summarise(mean_raw_Q = mean(raw_Q), .groups = "drop")  %>% 
  mutate( Z = ifelse(Z == 0, "Placebo", "Vaccine"))


p <- df_mark %>%
  mutate(
    raw_Q = (sequence_depth - K) / sequence_depth,
    Z = ifelse(Z == 0, "Placebo", "Vaccine"),
    y_order = rank(sequence_depth, ties.method = "first") 
  ) %>%
  filter(!is.na(raw_Q)) %>%
  ggplot(aes(x = raw_Q, y = y_order, color = Z, size = sqrt(sequence_depth))) +
  geom_quasirandom(
    width = 0,              # no horizontal spread
    groupOnX = FALSE,       # ensures jitter applied across all x values
    alpha = 0.8
  ) +
  geom_vline(xintercept = 0.99, linetype = "dashed", color = "red", size = 1.5) +
  geom_vline(data = means, aes(xintercept = mean_raw_Q, color = Z), size = 3) +
  facet_grid(rows = vars(Z), scales = "free_y") +
  scale_color_manual(values = c("Placebo" = "darkblue", "Vaccine" = "orange")) +
  xlab("Within-Host L832 Prevalence (Gaps Excluded)") +
  ylab(NULL) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_text(size = 30),
    axis.title = element_text(size = 30),
    strip.text = element_text(size = 30)
  )

ggsave(
  filename = "L832_prevalence_plot_gaps_excluded.pdf",
  plot = p,
  width = 16,       # inches
  height = 8,       # adjust as desired
  dpi = 600,        # publication-quality
  units = "in"
)

```



